Hypercoercions and a Framework for Equivalence of Cast Calculi

### 2019-10-25 paper.tex
minor edit
reformulation of statement of lemma 4.7
improve S~~D

%% Abstract
%% Note: \begin{abstract}...\end{abstract} environment must come
%% before \maketitle command
\begin{abstract}
  Designing a space-efficient cast representation that is good for
  both mechanized metatheory and implementation is
  challenging. Existing solutions are good for one or the other. This
  paper presents a new cast representation, named hypercoercions, that
  is good for both. On the way to proving the correctness of
  hypercoercions, this paper also makes progress on a general
  framework for proving the correctness of cast representations.
\end{abstract}

@@

\listoftodos{}

@@

\todo[inline]{I see lots of mistakes regarding when to make words
  plural (whether they should have an ``s'' at the end). Please review
  the rules of English regarding plurals. The rules are tricky...
  some words don't have a plural form, like ``blame''. -JS
  \\---\\
  Will fix grammar after fixing all other problems -KC}

@@

Subsection~\ref{sec:ld-hc} defines the $id$, $seq$, and $cast$
functions that construct
\lazyD\ hypercoercions. Subsection~\ref{sec:lud-hc} defines the
\lazyUD\ counterparts.  The functions that apply hypercoercions to
values are defined later in
Section~\ref{sec:hypercoercion-correctness}, after we introduce the
space-efficient machine and its definition of values in
Section~\ref{sec:framework}.

@@

Cast results ($r$) are the same as for \ineffCEK.

@@

The transition relation $\judgeSreduce{C}{s}{s}$ relies on functions provided 
by $C$ to work with casts.
%
When evaluating a cast expression, the machine extends the continuation with 
cast $cast(T_1,l,T_2)$. The function $ext(c,\kappa)$ composes $c$ with the cast 
at the top of $k$ by calling $seq$.
%
To construct a function, the machine fills the casts $id$.
%
To return a value to a continuation, the machine first applies the top cast to 
the value. If the application fails, the machine halts with the blame label 
from the failure. Otherwise, the machine handles the pre-continuation with 
$cont$.
%
When the machine constructs a pair, a left injection, or a right injection, it 
fills the casts with $id$ as well, just like how it did for functions. 
%
To apply a function, the machine applies the domain cast $c_1$ to the 
operand $v_2$, extends the continuation with the codomain cast $c_2$, and 
evaluate the function body.
%
To take out the first part of a pair, the machine returns the first value to a 
continuation extended with the first cast. Taking out the second part of a pair 
is similar.
%
%%In case splitting, if the target value is a left injection, the
%%machine moves to a state that will apply the first continuation
%%function to the value inside the left injection. The case for right
%%injection is similar.
%%
To case split a variant, the machine first look at the variant value. If the 
variant is a left injection, the machine composes the cast in the left 
injection with the domain casts of the first continuation function, then moves 
to a state that will apply the new function to the value in the left injection.
The case for right injections is similar.

Transitive closure of reduction ($\judgeSreduceTrans{C}{s}{s}$) and 
evaluation ($\judgeSeval{C}{e}{o}$) are standard. Value typing 
($\judgeType{v}{T}$) is straightforward.

@@

In this subsection, we prove that for all $C$, if $C$ is a Lazy D Cast
ADT, then
\[
  \evalEqv{\effCEK{C}}{\ineffCEKD}
\]
The main work of the proof is in a lemma that establishes a weak
bisimulation between \effCEK{C} and \ineffCEKD{}.  The bisimulation
goes more smoothly if we require two more properties of the cast
representation: that the casts form a monoid (the $id$ and $seq$
operators) and that the cast constructor is equivalent to the identity
operator when applied to an identical source and target type.
However, the final theorem does not require these two properties
because we can prove 1) that \effCEK{C_1} and \effCEK{C_2} are
equivalent for any two Lazy D Cast ADTs $C_1$ and $C_2$
(Proposition~\ref{thm:surely-lazyD-eqv}), and 2) there is an instance
of the Lazy D Cast ADT, named $L$, that is a monoid and satisfies the
two extra properties, and is therefore equivalent to \ineffCEKD{}
(Lemma~\ref{lem:L-correct}). So by transitivity, for any Lazy D Cast
ADT $C$ we have \evalEqv{\effCEK{C}}{\ineffCEKD}.

@@

We shall prove that if a instance of Cast ADT $C$ is Lazy D, is a
monoid, and if $cast(T,l,T) = id(T)$, then there is
%% $eval_\mathcal{D} = eval_{\mathcal{S}(C)}$. We proof this lemma with
a weak bisimulation between \effCEK{C} and \ineffCEKD{}.
We define monoid as follows.

@@

\todo[inline]{How does the following bisimulation relation
  compare to the one used in the Blame and Coercion paper?
  Are there any guiding principles to how this bisimulation
  relation was constructed? -JS
\\---\\
The guiding principle is to keep the change in D side atomic. \\
In the blame and coercion paper,
the bisimulation relation between $ \lambda{}C $ and $ \lambda{}S $ seems to be 
related, but its (iii) rules in Figure 6 looks wired to me. -KC}

Fig.~\ref{fig:bisim-SC-D} defines the bisimulation relation.  It is
the smallest congruence relation that also includes the pairs induced
by the rules in Fig.~\ref{fig:bisim-SC-D}.
%
The relation is mutually defined on continuations, written $\kappa
\approx \kappa$, and values, written $v \approx v : T$, where $T$ is
their type.
%
If $k \approx \kappa$, adding an identity cast on top of $k$ gives a related 
continuation.
%
\todo[inline]{The use of ``wrap'' obscures this explanation.
  Consider getting rid of wrap by inlining it. -JS
\\---\\
I agree that inlining wrap and ext can improve readability of $\kappa \approx 
\kappa$, but this would make latex different from Agda, is it alright?
-KC
\\---\\
Yes, that is alright. Inlining is really a trivial change.
-JS
}
%
If $\kappa_1 \approx \kappa_2$, extending $\kappa_1$ and $\kappa_2$ with the 
same cast gives related results.
%
If two values of type $P$ are related, injecting them to $\TOOdyn$ gives 
related values.
%
% The following is part of being a congruence. -JS
%Unit values are of course related. 
%
To relate to a naked function in \ineffCEKD, the function in \effCEK{C} 
must use identities as its casts.
%
If functions are related, we can apply the same cast on both side.
%
Rules for products and sums are similar.

@@

  \fbox{$\kappa \approx \kappa$}
  \begin{gather*}
  \inference{
    k \approx \kappa
  }{
    wrap(k) \approx \kappa
  }
  \quad
  \inference{
    \kappa_1 \approx \kappa_2
  }{
    ext(cast(T_1,l,T_2),\kappa_1) \approx 
    \kOOcast{\cOOcast{T_1}{l}{T_2}}{\kappa_2}
  }
\\
  \end{gather*}
\\  
  \fbox{$v \approx v : T$}
  \begin{gather*}
  \inference{
    v_1 \approx v_2 : P
  }{
    \hcvOOinj{P}{v_1} \approx \vOOcast{v_2}{\cOOcast{P}{l}{\TOOdyn}}
    : \TOOdyn
  }
\\ \\
   \inference{}{
    \hcvOOtt \approx \vOOtt : \POOunit
   }
   \quad
  \inference{}{
    \hcvOOtt \approx \vOOcast{\vOOtt}{\cOOcast{\POOunit}{l}{\POOunit}}
    : \POOunit
  }
  \\ \\
  \inference{
    \mathcal{E}_1 \approx \mathcal{E}_2 &
    dom(\mathcal{E}_1) = dom(\mathcal{E}_2) = \Gamma \\
    \Gamma , x : T_1 \vdash e : T_2
  }{
    \hcvOOfun{id(T_1)}{\mathcal{E}_1}{x}{e}{id(T_2)}
    \approx
    \vOOfun{\mathcal{E}_2}{x}{e}
    : \POOfun{T_1}{T_2}
  }
\\ \\
\inference{
\hcvOOfun{c_1}{\mathcal{E}}{x}{e}{c_2} \approx v_2 : \POOfun{T_1}{T_2} \\
c_1' = seq(cast(T_3,l,T_1),c_1) \\
c_2' = seq(c_2,cast(T_2,l,T_4))
}{
\hcvOOfun{c_1'}{\mathcal{E}}{x}{e}{c_2'}
\approx
\vOOcast{v_2}{\cOOcast{\POOfun{T_1}{T_2}}{l}{\POOfun{T_3}{T_4}}}
: \POOfun{T_3}{T_4}
}
\\ \\
  \inference{
    v_1 \approx v_2 : T_1 &
    v_3 \approx v_4 : T_2 &
  }{
    \hcvOOcons{v_1}{id(T_1)}{v_3}{id(T_2)}
    \approx
    \vOOcons{v_2}{v_4}
    : \POOprod{T_1}{T_2}
  }
\\ \\
  \inference{
    \hcvOOcons{v_1}{c_1}{v_3}{c_2} \approx v_2 : \POOprod{T_1}{T_2} \\
    c_1' = seq(c_1,cast(T_1,l,T_3)) \\
    c_2' = seq(c_2,cast(T_2,l,T_4))
  }{
    \hcvOOcons{v_1}{c_1'}{v_3}{c_2'} \approx
    \vOOcast{v_2}{\cOOcast{\POOprod{T_1}{T_2}}{l}{\POOprod{T_3}{T_4}}}
    : \POOprod{T_3}{T_4}
  }
\\ \\
  \inference{
    v_1 \approx v_2 : T_1
  }{
    \hcvOOinl{v_1}{id(T_1)} \approx \vOOinl{v_2}
    : \POOsum{T_1}{T_2}
  }
  \quad
  \inference{
    \hcvOOinl{v_1}{c} \approx v_2
    : \POOsum{T_1}{T_2} &
    c' = seq(c,cast(T_1,l,T_3))
  }{
    \hcvOOinl{v_1}{c'} \approx
    \vOOcast{v_2}{\cOOcast{\POOsum{T_1}{T_2}}{l}{\POOsum{T_3}{T_4}}}
    : \POOsum{T_3}{T_4} 
  }
  \\
  \inference{
  v_1 \approx v_2 : T_2
  }{
  \hcvOOinr{v_1}{id(T_2)} \approx \vOOinr{v_2}
  : \POOsum{T_1}{T_2}
  }
  \quad
  \inference{
  \hcvOOinr{v_1}{c} \approx v_2
  : \POOsum{T_1}{T_2} &
  c' = seq(c,cast(T_2,l,T_4))
  }{
  \hcvOOinr{v_1}{c'} \approx
  \vOOcast{v_2}{\cOOcast{\POOsum{T_1}{T_2}}{l}{\POOsum{T_3}{T_4}}}
  : \POOsum{T_3}{T_4}
  }
  \end{gather*}
  \caption{Bisimulation between \effCEK{C} and \ineffCEKD}
  \label{fig:bisim-SC-D}
\end{figure}

\todo[inline]{Give the motivation for the following lemma. Where
  does it get used? As in, which case of which other lemma or theorem? -JS
\\---\\
Partially fixed. Will say in which cases after finishing the bisim proof. -KC}

We use the following lemma in our bisimulation proof.

\begin{lemma}[SoAndSo]
  \label{thm:lem-apply-cast-SD}
  Assume $C$ implements Lazy D Cast ADT
  and $v_1$ is a value in \effCEK{C}
  and $v_2$ is a value in \ineffCEKD\
  and $v_1 \approx v_2 : T_1$,
  \[
  applyCast_{\text{\effCEK{C}}}(v,cast(T_1,l,T_2)) 
  \approx 
  applyCast_{\text{\ineffCEKD}}(v,\cOOcast{T_1}{l}{T_2})
  \]
\end{lemma}
\begin{proof}
  Case splitting whether $T_1$ is shallowly consistent with $T_2$.
  If yes, case splitting how they are shallowly consistent.
  Then apply properties (3) - (11) of Lazy D Cast when applicable.
\end{proof}

Before proving the weak bisimulation between \effCEK{C} and \ineffCEKD, we 
introduce one more definition.

@@

\todo[inline]{Explain that this is an unusual formulation of a weak
  bisimulation.  We promise it will all be OK regarding how it is used
  in the correctness proof for S(C). -JS}

\begin{lemma}[Weak Bisimulation between \effCEK{C} and \ineffCEKD]
  \label{thm:surely-monoidic-reduce}
  Assume $C$ implements Lazy D Cast ADT
  and $C$ is a monoid
  and $cast(T,l,T)=id(T)$,
  $ s_{1} \in \text{\effCEK{C}}$,
  $ s_{2} \in \text{\ineffCEKD}$,
  $ s_{1} \approx s_{2}$,
\begin{enumerate}
  \item If $s_1 = \sOOhalt{o}$ then $s_2 = \sOOhalt{o}$
  \item If $s_2 = \sOOhalt{o}$ then $s_1 = \sOOhalt{o}$
  \item If $\judgeSreduce{C}{s_1}{s_3}$ 
    then $\judgeDreduce{s_2}{s_4}$ for some $s_4$
  \item If $\judgeDreduce{s_2}{s_4}$
    then $\judgeSreduce{C}{s_1}{s_3}$ for some $s_3$
  \item If $\judgeSreduce{C}{s_1}{s_3}$ 
    and  $\judgeDreduce{s_2}{s_4}$ 
    then $s_3 \approx^{*} s_4$
\end{enumerate}
  Or diagrammatically, 
  \[\begin{array}{clclc}
  s_{1} & \longrightarrow_{\text{\effCEK{C}}} & s_1'\\
  \rotatebox[origin=c]{90}{$\approx$} 
  & & & \;\;\;\rotatebox[origin=c]{-20}{$\approx$} \\
  s_{2} & 
  \longrightarrow_{\text{\ineffCEKD}} & s_2' &
  \longrightarrow^{*}_{\text{\ineffCEKD}} & s_2'' \\
  \end{array}\]
\end{lemma}
\begin{proof}
  Part (1) and (2) follow from the definition of $s \approx s$.
  Part (3) and (4) are true because we have transition rules for every 
  non-halting state.
  Part (5) is more interesting. Our goal is to prove that one reduction step 
  in  \effCEK{C} corresponds to one or more steps in \ineffCEKD.  
  Consider
  \[
  \sOOreturn{\hcvOOcons{v_{11}}{c_{11}}{v_{12}}{c_{12}}}{
    wrap(\kOOcar{\kappa})
  }
  \approx
  \sOOreturn{v_2}{\kOOcar{\kappa}}
  \]

  By looking at the definition of $v \approx v$, we know
  \[ c_{11} = seq(\dots seq(id(T_1),cast(T_1,l,T_3)) \dots) \]
  \[ c_{12} = seq(\dots seq(id(T_2),cast(T_2,l,T_4)) \dots) \] 
  \[ v_2 = \vOOcast{(\vOOcons{v_{21}}{v_{22}})}{
    \cOOcast{\POOprod{T_1}{T_2}}{l_1}{\POOprod{T_3}{T_4}}} \dots
  \]
  Thus we have the following reduction in \effCEK{C}. The equation 
  reasoning employs the fact that $C$ is a monoid.
  \[
  \begin{array}{ll}
&
\sOOreturn{\hcvOOcons{v_{11}}{c_{11}}{v_{12}}{c_{12}}}{wrap(\kOOcar{\kappa})}
\\
\longmapsto_\text{\effCEK{C}} &
\sOOreturn{v_{11}}{ext(c_{11},\kappa)}
\\
= &
\sOOreturn{v_{11}}{ext(seq(\dots seq(id(T_1),cast(T_1,l,T_3))\dots),\kappa)}
\\
= &
\sOOreturn{v_{11}}{ext(cast(T_1,l,T_3),\dots,\kappa)}
  \end{array}
  \]

And we have the following reduction in \ineffCEKD, where the last state is 
related to the last state in the above sequence.

@@

Other cases are similar. Of course we need the lemma about $applyCast$ 
(\ref{thm:lem-apply-cast-SD}) when cast applications happen in both machines. 
And in case spliting, we shall use $cast(T,l,T)=id(T)$ to relate the 
continuation functions.
\end{proof}

\todo[inline]{Should I invest time on restating the bisimulation lemma in the 
standard format? -KC}

\begin{corollary}[Correctness of \effCEK{C}]
  \label{thm:surely-monoidic-eval}
  Suppose $C$ is an instance of the Cast ADT
  and the \lazyD{} ADT, $C$ is a monoid,
  and $cast(T,l,T)=id(T)$.
  If $\judgetype{\emptyset}{e}{T}$ and $o : T$, then
  \[
    \text{\evalEqv{\effCEK{C}}{\ineffCEKD}}
  \]
\end{corollary}
\begin{proof}
  Proving that the left-hand side implies the right-hand side is
  trivial.  To prove $\judgeDeval{e}{o}$ implies
  $\judgeSeval{C}{e}{o}$, we first take one step in \effCEK{C}, and
  catch it up in \ineffCEKD\ by taking one or more steps in
  \ineffCEKD. As we just took a positive number of steps, we are
  closer to the halting state.
  \todo[inline]{I might want to revise this proof. -KC}
\end{proof}

@@

%% In this subsection, we prove that for all $C$, if $C$ is a Lazy D Cast ADT, 
%% then \evalEqv{\ineffCEKD}{\effCEK{C}}. We first 
%% prove that all $S(C)$ where $C$ is a Lazy D Cast ADT are equivalent, then 
%% connect this theorem to Lemma~\ref{thm:surely-monoidic-eval}.

@@

We prove the equivalence among two \effCEK{C} machines with a strong
bisimulation. The bisimulation relation is the smallest congruence relation 
that also includes the pairs induced by the rules below.

\begin{gather*}
\inference{
}{
  cast_1(T_1,l,T_2) \approx cast_2(T_1,l,T_2)
}
\quad
\inference{
}{
  id_1(T) \approx id_2(T)
}
\quad
\inference{
  c_1 \approx c_2 &
  c_3 \approx c_4
}{
  seq_1(c_1,c_3) \approx seq_2(c_2,c_4)
}
\end{gather*}

>>>>

%% Abstract
%% Note: \begin{abstract}...\end{abstract} environment must come
%% before \maketitle command
\begin{abstract}
  Designing a space-efficient cast representation that is good for
  both mechanized metatheory and implementation is
  challenging. Existing solutions (i.e. coercions, threesomes, and 
  supercoercions) are good for one or the other. This
  paper presents a new cast representation, named hypercoercions, that
  is good for both. On the way to proving the correctness of
  hypercoercions, this paper also makes progress on a general
  framework for proving the correctness of cast representations.
\end{abstract}

@@

@@

\todo[inline]{
	Check grammar;
	Improve the proof of \ref{thm:SD-bisim}.
	And the proof of \ref{thm:SD-equiv}
}

@@

Subsection~\ref{sec:ld-hc} defines the $id$, $seq$, and $cast$
functions that construct
\lazyD\ hypercoercions. Subsection~\ref{sec:lud-hc} defines the
\lazyUD\ counterparts.  The functions that apply hypercoercions to
values are defined later in
Section~\ref{sec:hypercoercion-correctness}, after we introduce the framework 
and its new space-efficient definition of values in Section~\ref{sec:framework}.

@@

Cast results ($r$) are the same as for \ineffCEK. 
Value typing ($\judgeType{v}{T}$) is defined in the obvious way.

@@

The transition relation $\judgeSreduce{C}{s}{s}$ relies on functions provided 
by $C$ to work with casts.
%
When evaluating a cast expression, the machine extends the continuation with 
cast $cast(T_1,l,T_2)$. The function $ext(c,\kappa)$ composes $c$ with the cast 
at the top of $k$ by calling $seq$.
%
To construct a function, the machine fills the casts with $id$s.
%
To return a value to a continuation, the machine first applies the top cast to 
the value. If the application fails, the machine halts with the blame label 
from the failure. Otherwise, the machine handles the pre-continuation with 
$cont$.
%
When the machine constructs a pair, a left injection, or a right injection, it 
fills the casts with $id$ as well, just like how it did for functions. 
%
To apply a function, the machine applies the domain cast $c_1$ to the 
operand $v_2$, extends the continuation with the codomain cast $c_2$, and 
evaluate the function body.
%
To take out the first part of a pair, the machine returns the first value to a 
continuation extended with the first cast. Taking out the second 
part of a pair is similar.
%
%%In case splitting, if the target value is a left injection, the
%%machine moves to a state that will apply the first continuation
%%function to the value inside the left injection. The case for right
%%injection is similar.
%%
To case split a variant, the machine first look at the variant value. If the 
variant is a left injection, the machine composes the cast in the left 
injection with the domain cast of the first continuation function, then move 
to a state that will apply the new function to the value in the left injection.
The case for right injections is similar.

Reflexive transitive closure of reduction ($\judgeSreduceTrans{C}{s}{s}$) and
evaluation ($\judgeSeval{C}{e}{o}$) are the same as for \ineffCEK.

@@

In this section, we prove that for all $C$, if $C$ is a Lazy D Cast
ADT, then
\[
  \evalEqv{\effCEK{C}}{\ineffCEKD}
\]
The main work of the proof is in a lemma that establishes a weak
bisimulation between \effCEK{C} and \ineffCEKD{}.  The bisimulation
goes more smoothly if we require two more properties of the cast
representation: that the casts form a monoid (the $id$ and $seq$
operators) and that the cast constructor is equivalent to the identity
operator when applied to an identical source and target type.
However, the final theorem does not require these two properties
because we can prove 1) that \effCEK{C_1} and \effCEK{C_2} are
equivalent for any two Lazy D Cast ADTs $C_1$ and $C_2$
(Proposition~\ref{thm:surely-lazyD-eqv}), and 2) there is an instance
of the Lazy D Cast ADT, named $L$, that is a monoid and satisfies the
two extra properties, and is therefore equivalent to \ineffCEKD{}
(Lemma~\ref{lem:L-correct}). So by transitivity, for any Lazy D Cast
ADT $C$ we have \evalEqv{\effCEK{C}}{\ineffCEKD}.

@@

We shall prove that if an instance of Cast ADT $C$ is Lazy D, is a
monoid, and if $cast(T,l,T) = id(T)$, then there is
%% $eval_\mathcal{D} = eval_{\mathcal{S}(C)}$. We proof this lemma with
a weak bisimulation between \effCEK{C} and \ineffCEKD{}.
We define monoid as follows.

@@

%\todo[inline]{How does the following bisimulation relation
%  compare to the one used in the Blame and Coercion paper?
%  Are there any guiding principles to how this bisimulation
%  relation was constructed? -JS
%\\---\\
%The guiding principle is to keep the change in D side atomic. \\
%In the blame and coercion paper,
%the bisimulation relation between $ \lambda{}C $ and $ \lambda{}S $ seems to 
%be 
%related, but its (iii) rules in Figure 6 looks wired to me. -KC}

Fig.~\ref{fig:bisim-SC-D} defines the bisimulation relation.  It is
the smallest congruence relation that also includes the pairs induced
by the rules in Fig.~\ref{fig:bisim-SC-D}.
%
The relation is mutually defined on continuations, written $\kappa
\approx \kappa : T$, where $T$ is their source type, and values, written $v 
\approx v : T$, where $T$ is their type.
%
If $k \approx \kappa : T$, adding an identity cast on top of $k$ gives a 
related continuation.
%
%\todo[inline]{The use of ``wrap'' obscures this explanation.
%  Consider getting rid of wrap by inlining it. -JS
%\\---\\
%I agree that inlining wrap and ext can improve readability of $\kappa \approx 
%\kappa$, but this would make latex different from Agda, is it alright?
%-KC
%\\---\\
%Yes, that is alright. Inlining is really a trivial change.
%-JS
%}
%
If $\kOOcast{c}{k_1} \approx \kappa_2 : T$, extending them with 
the same cast gives related continuations.
%
If two values of type $P$ are related, injecting them to $\TOOdyn$ gives 
related values.
%
% The following is part of being a congruence. -JS
%Unit values are of course related. 
%
To relate to a naked function in \ineffCEKD, the function in \effCEK{C} 
must use identities as its casts.
%
If functions are related, we can apply the same cast on both side.
%
Rules for products and sums are similar.

@@

  \fbox{$\kappa \approx \kappa : T$}
  \begin{gather*}
  \inference{
    k \approx \kappa : T
  }{
    \kOOcast{id(T)}{k} \approx \kappa
  }
  \quad
  \inference{
    \kOOcast{c}{k_1} \approx \kappa_2 : T_2
  }{
    \kOOcast{seq(cast(T_1,l,T_2),c)}{k_1} \approx 
    \kOOcast{\cOOcast{T_1}{l}{T_2}}{\kappa_2}
  }
\\
  \end{gather*}
\\  
  \fbox{$v \approx v : T$}
  \begin{gather*}
  \inference{
    v_1 \approx v_2 : P
  }{
    \hcvOOinj{P}{v_1} \approx \vOOcast{v_2}{\cOOcast{P}{l}{\TOOdyn}}
    : \TOOdyn
  }
\\ \\
   \inference{}{
    \hcvOOtt \approx \vOOtt : \POOunit
   }
   \quad
  \inference{}{
    \hcvOOtt \approx \vOOcast{\vOOtt}{\cOOcast{\POOunit}{l}{\POOunit}}
    : \POOunit
  }
  \\ \\
  \inference{
    \mathcal{E}_1 \approx \mathcal{E}_2 &
    dom(\mathcal{E}_1) = dom(\mathcal{E}_2) = \Gamma \\
    \Gamma , x : T_1 \vdash e : T_2
  }{
    \hcvOOfun{id(T_1)}{\mathcal{E}_1}{x}{e}{id(T_2)}
    \approx
    \vOOfun{\mathcal{E}_2}{x}{e}
    : \POOfun{T_1}{T_2}
  }
\\ \\
\inference{
\hcvOOfun{c_1}{\mathcal{E}}{x}{e}{c_2} \approx v_2 : \POOfun{T_1}{T_2} \\
c_1' = seq(cast(T_3,l,T_1),c_1) \\
c_2' = seq(c_2,cast(T_2,l,T_4))
}{
\hcvOOfun{c_1'}{\mathcal{E}}{x}{e}{c_2'}
\approx
\vOOcast{v_2}{\cOOcast{\POOfun{T_1}{T_2}}{l}{\POOfun{T_3}{T_4}}}
: \POOfun{T_3}{T_4}
}
\\ \\
  \inference{
    v_1 \approx v_2 : T_1 &
    v_3 \approx v_4 : T_2 &
  }{
    \hcvOOcons{v_1}{id(T_1)}{v_3}{id(T_2)}
    \approx
    \vOOcons{v_2}{v_4}
    : \POOprod{T_1}{T_2}
  }
\\ \\
  \inference{
    \hcvOOcons{v_1}{c_1}{v_3}{c_2} \approx v_2 : \POOprod{T_1}{T_2} \\
    c_1' = seq(c_1,cast(T_1,l,T_3)) \\
    c_2' = seq(c_2,cast(T_2,l,T_4))
  }{
    \hcvOOcons{v_1}{c_1'}{v_3}{c_2'} \approx
    \vOOcast{v_2}{\cOOcast{\POOprod{T_1}{T_2}}{l}{\POOprod{T_3}{T_4}}}
    : \POOprod{T_3}{T_4}
  }
\\ \\
  \inference{
    v_1 \approx v_2 : T_1
  }{
    \hcvOOinl{v_1}{id(T_1)} \approx \vOOinl{v_2}
    : \POOsum{T_1}{T_2}
  }
  \quad
  \inference{
    \hcvOOinl{v_1}{c} \approx v_2
    : \POOsum{T_1}{T_2} &
    c' = seq(c,cast(T_1,l,T_3))
  }{
    \hcvOOinl{v_1}{c'} \approx
    \vOOcast{v_2}{\cOOcast{\POOsum{T_1}{T_2}}{l}{\POOsum{T_3}{T_4}}}
    : \POOsum{T_3}{T_4} 
  }
  \\
  \inference{
  v_1 \approx v_2 : T_2
  }{
  \hcvOOinr{v_1}{id(T_2)} \approx \vOOinr{v_2}
  : \POOsum{T_1}{T_2}
  }
  \quad
  \inference{
  \hcvOOinr{v_1}{c} \approx v_2
  : \POOsum{T_1}{T_2} &
  c' = seq(c,cast(T_2,l,T_4))
  }{
  \hcvOOinr{v_1}{c'} \approx
  \vOOcast{v_2}{\cOOcast{\POOsum{T_1}{T_2}}{l}{\POOsum{T_3}{T_4}}}
  : \POOsum{T_3}{T_4}
  }
  \end{gather*}
  \caption{Bisimulation between \effCEK{C} and \ineffCEKD}
  \label{fig:bisim-SC-D}
\end{figure}

%\todo[inline]{Give the motivation for the following lemma. Where
%  does it get used? As in, which case of which other lemma or theorem? -JS
%\\---\\
%Partially fixed. Will say in which cases after finishing the bisim proof. -KC}

We use the following lemma in our bisimulation proof when cast applications 
happen.

\begin{lemma}[$applyCast$ Preserves Bisimulation]
  \label{thm:lem-apply-cast-SD}
  Assume $C$ implements Lazy D Cast ADT,
  $v_1$ is a value in \effCEK{C},
  $v_2$ is a value in \ineffCEKD,
  and $v_1 \approx v_2 : T_1$,
  \[
  applyCast_{\text{\effCEK{C}}}(v_1,cast(T_1,l,T_2)) 
  \approx 
  applyCast_{\text{\ineffCEKD}}(v_2,\cOOcast{T_1}{l}{T_2})
  \]
\end{lemma}
\begin{proof}
  Case splitting whether $T_1$ is shallowly consistent with $T_2$.
  If yes, case splitting how they are shallowly consistent.
  Then apply properties (3) - (11) of Lazy D Cast when applicable.
\end{proof}

Before proving the weak bisimulation between \effCEK{C} and \ineffCEKD, we 
introduce one more definition. $s_1 \approx^{*} s_2$ means we will get $s_1 
\approx s_2$ after more transition steps in \ineffCEKD.

@@

%\todo[inline]{Explain that this is an unusual formulation of a weak
%  bisimulation.  We promise it will all be OK regarding how it is used
%  in the correctness proof for S(C). -JS}

The lemma below formulate the weak bisimulation between \effCEK{C} and 
\ineffCEKD. It is unusual because we use $s \approx s$ in premise while $s 
\approx^{*} s$ in conclusion. As we will show in the correctness proof for 
\effCEK{C}, however, this formulation is fine for our purpose.

\begin{lemma}[Weak Bisimulation between \effCEK{C} and \ineffCEKD]
  \label{thm:SD-bisim}
  Assume $C$ implements Lazy D Cast ADT
  and $C$ is a monoid
  and $cast(T,l,T)=id(T)$,
  $s_{1} \in \text{\effCEK{C}}$,
  $s_{2} \in \text{\ineffCEKD}$,
  $s_{1} \approx s_{2}$,
\begin{enumerate}
  \item If $s_1 = \sOOhalt{o}$ then $s_2 = \sOOhalt{o}$
  \item If $s_2 = \sOOhalt{o}$ then $s_1 = \sOOhalt{o}$
  \item If $\judgeSreduce{C}{s_1}{s_3}$ 
    then $\judgeDreduce{s_2}{s_4}$ 
    and  $s_3 \approx^{*} s_4$
    for some $s_4$
  \item If $\judgeDreduce{s_2}{s_4}$
    then $\judgeSreduce{C}{s_1}{s_3}$ 
    and  $s_3 \approx^{*} s_4$
    for some $s_3$
\end{enumerate}
  Or diagrammatically, 
  \[\begin{array}{clclc}
  s_{1} & \longrightarrow_{\text{\effCEK{C}}} & s_1'\\
  \rotatebox[origin=c]{90}{$\approx$} 
  & & & \;\;\;\rotatebox[origin=c]{-20}{$\approx$} \\
  s_{2} & 
  \longrightarrow_{\text{\ineffCEKD}} & s_2' &
  \longrightarrow^{*}_{\text{\ineffCEKD}} & s_2'' \\
  \end{array}\]
\end{lemma}
\begin{proof}
  Part (1) and (2) follow from the definition of $s \approx s$.
  Part (3) and (4) are more interesting. 
  By the definition of $s \approx s$, it is straightforward to prove 
  $\judgeSreduce{C}{s_1}{s_3}$ if and only if $\judgeDreduce{s_2}{s_4}$.
  So we can prove two parts together by showing that 
  \[
  \text{if}\; s_{1} \approx s_{2}
  \;\text{and}\;\judgeSreduce{C}{s_1}{s_3}
  \;\text{and}\;\judgeDreduce{s_2}{s_4}
  \;\text{then}\; s_3 \approx^{*} s_4.
  \]
  To put it another way, our goal is to proving 
  one reduction step in \effCEK{C} corresponds to
  one or more steps in \ineffCEKD.  
  Although there are many cases, the general idea is the same.
  Consider
  \[
  \sOOreturn{\hcvOOcons{v_{11}}{c_{11}}{v_{12}}{c_{12}}}{
    wrap(\kOOcar{\kappa})
  }
  \approx
  \sOOreturn{v_2}{\kOOcar{\kappa}}
  \]

  By looking at the definition of $v \approx v$, we know
  \[ c_{11} = seq(seq(id(T_1),cast(T_1,l,T_3)) \dots) \]
  \[ c_{12} = seq(seq(id(T_2),cast(T_2,l,T_4)) \dots) \] 
  \[ v_2 = \vOOcast{(\vOOcons{v_{21}}{v_{22}})}{
    \cOOcast{\POOprod{T_1}{T_2}}{l_1}{\POOprod{T_3}{T_4}}} \dots
  \]
  Thus we have the following reduction in \effCEK{C}. The equation 
  reasoning employs the premise that $C$ is a monoid.
  \[
  \begin{array}{ll}
&
\sOOreturn{\hcvOOcons{v_{11}}{c_{11}}{v_{12}}{c_{12}}}{wrap(\kOOcar{\kappa})}
\\
\longmapsto_\text{\effCEK{C}} &
\sOOreturn{v_{11}}{ext(c_{11},\kappa)}
\\
= &
\sOOreturn{v_{11}}{ext(seq(seq(id(T_1),cast(T_1,l,T_3))\dots),\kappa)}
\\
= &
\sOOreturn{v_{11}}{
	ext(id(T_1),ext(cast(T_1,l,T_3),ext(\dots,\kappa)))
}
\\
= &
\sOOreturn{v_{11}}{ext(cast(T_1,l,T_3),ext(\dots,\kappa)}
  \end{array}
  \]

And we have the following reduction in \ineffCEKD, where the last state is 
related to the above \effCEK{C} state.

@@

Other cases are similar. Of course we need the lemma about $applyCast$ 
(\ref{thm:lem-apply-cast-SD}) when cast applications happen in both machines. 
And in case spliting, we employ $cast(T,l,T)=id(T)$ to relate the continuation 
functions.
\end{proof}

\begin{corollary}[Correctness of \effCEK{C}]
  \label{thm:SD-equiv}
  Suppose $C$ is an instance of the Cast ADT
  and the \lazyD{} ADT, $C$ is a monoid,
  and $cast(T,l,T)=id(T)$.
  If $\judgetype{\emptyset}{e}{T}$ and $o : T$, then
  \[
    \text{\evalEqv{\effCEK{C}}{\ineffCEKD}}
  \]
\end{corollary}
\begin{proof}
  Proving that the left-hand side implies the right-hand side is
  trivial.  To prove $\judgeDeval{e}{o}$ implies
  $\judgeSeval{C}{e}{o}$, we first take one step in \effCEK{C}, and
  catch it up in \ineffCEKD\ by taking one or more steps in
  \ineffCEKD. As we just took a positive number of steps, we are
  closer to the halting state.
\end{proof}

@@

%% In this subsection, we prove that for all $C$, if $C$ is a Lazy D Cast ADT, 
%% then \evalEqv{\ineffCEKD}{\effCEK{C}}. We first 
%% prove that all $S(C)$ where $C$ is a Lazy D Cast ADT are equivalent, then 
%% connect this theorem to Lemma~\ref{thm:SD-equiv}.

@@

We prove the equivalence among two \effCEK{C} machines with a strong
bisimulation. The bisimulation relation is the smallest congruence relation 
that also includes the rules below.
\begin{gather*}
\inference{
}{
  cast_1(T_1,l,T_2) \approx cast_2(T_1,l,T_2)
}
\quad
\inference{
}{
  id_1(T) \approx id_2(T)
}
\quad
\inference{
  c_1 \approx c_2 &
  c_3 \approx c_4
}{
  seq_1(c_1,c_3) \approx seq_2(c_2,c_4)
}
\end{gather*}
